The Galactic Midplane Is Not a Plane: Implications for Dynamical Analysis with {\em Gaia} Data and Beyond

### 2019-05-04 paper/ms.tex
Add discussion section on coordinate system offsets

Importantly, the location of \sgra{} may not be equivalent to
the location of the ``dynamical Galactic center,'' i.e., the point in
three-dimensional space about which the stars in the solar neighborhood are
orbiting. This assumption, although sensible and frequently made, has not yet been
justified. It is tested in any construction of a  dynamical model where the
distance from the Sun to the Galactic center, $R_0$, is a free parameter. For
example, \citet{2015ApJ...803...80K} measured $R_0$ while modelling the
dynamics of the stream Palomar 5. Many other dynamical measurements of $R_0$
have been made (\citealt{2016ARAA..54..529B} summarize many pre-{\em Gaia}
results) but none have yet achieved a precision comparable to that of the
distance to \sgra{}.

@@

Galactocentric coordinate system offsets can also be generated by errors in
the parameters of the Galactocentric coordinate system. However, such
parameters have been measured to high enough precision that they are likely
subdominant to the intrinsic variation of the midplane across large distances.
On shorter scales, they are likely the dominant source of error, and they are
likely to be important for detailled modelling of kinematically cold stars.

@@

>>>>

@@

@@

\subsection{Coordinate System Offsets}\label{ssec:coord_off}
In Section~\ref{sec:mes_err} we discussed the current observational efforts to
measure the parameters of the Galactocentric coordinate system. We noted that
the solar height is discrepant between gas-based and stellar-based
determinations by $\sim5\,\pc$. While the effect is small and thus only likely
to dominate the intrinsic midplane variations on small scales, it may be
important for detailed modeling of kinematically cold stars where a $5\,\pc$
offset is relevant.

We also noted that the distance from the Sun to \sgra{} has been recently
measured to a very low uncertainty of $\sim0.3\%$ or $\pm25\,\pc$
\citep{2019arXiv190405721A}. However, the location of \sgra{} may not be
equivalent to the location of the ``dynamical Galactic center,'' i.e., the
point in three-dimensional space about which the stars in the solar
neighborhood are orbiting. This assumption, although sensible and frequently
made, has not yet been justified.

If the dynamical Galactic center is offset from \sgra{} by $100\,\pc$, only a
$1.2\%$ difference, then this induces a $21\%$ error in $J_R$ (see
Section~\ref{ssec:quant}). The reason such a large error in $J_R$ can be
generated by a small error in $R_0$ can be understood from the epicyclic
approximation, which states that $\Delta J_R/J_R = 2\Delta R/A_R$. The fractional
error in $J_R$ is related to the error in $R_0$ as a fraction of the {\em
radial amplitude} of the orbit, which is much smaller than $R_0$
($\sim1.2\,\kpc$ for the thin-disk and thick-disk orbits we considered). This
also implies the very precise $0.3\%$ measurement of $R_0$ still translates to
a $\sim4\%$ uncertainty in $J_R$.

The assumption that the dynamical Galactic center and \sgra{} are colocated is
tested in any construction of a dynamical model where $R_0$ is a free
parameter. For example, \citet{2015ApJ...803...80K} measured $R_0$ while
modeling the dynamics of the stream Palomar 5. Many other dynamical
measurements of $R_0$ have been made (\citealt{2016ARAA..54..529B} summarize
many pre-{\em Gaia} results), but none have yet achieved a precision
comparable to that of the distance to \sgra{}.

We did not consider in this text the effect of the {\em angular position} of
the dynamical Galactic center being offset with \sgra{}.
